---
title: "Midterm Review"
author: "Reese Quillian"
date: "2022-10-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
library(psych)
library(ggplot2)
library(GGally)
library(devtools)
library(ggfortify)
library(here)
library(dplyr)
library(ggpubr)
library(MASS)
library(lindia)

setwd("C:/Users/Student/OneDrive - University of Virginia/Documents/SYS4021/")

# read in data
housing <- read.csv("housing.csv")
```


### Question 1 of 19

Use a box plot to identify if the price variable has any outliers. If, so how many does it have? ____________

```{r}
ggplot(data=housing, aes(x=price)) + geom_boxplot()
# no outliers
```


### Question 2 of 19

Use a box plot to determine the value of the upper whisker of Price. What is this value? ___________

```{r}
summary(housing$price)
boxplot.stats(housing$price)
# upper whisker is max, so 1250000
```


### Question 3 of 19

Use a scatter plot matrix to identify the variable in the data set that has the strongest linear relationship with price.

 **A. sqft**
 B. City
 C. bedrooms
 D. baths

```{r}
pairs.panels(housing[,c("sqft", "City", "bedrooms", "baths", "price")])
```


### Question 4 of 19

Use a scatter plot matrix to identify the correlation coefficient between price and the variable in the data set that has the strongest linear relationship with price. _______________

```{r}
# strongest correlation is 0.65 
# price and sqft
```
 

### Question 5 of 19

Build a main effects model, house.main, using sqft, bedrooms, and baths. The response variable is price. 

Which of the following statement(s) are true?

 **A. The model is significant at the p<0.01 level.**	
 B. All of the predictors are significant at the p<0.01 level	
 C. None of the predictors are significant at the p<0.01 level.	
 **D. Only one of the coefficients is significant at the p<0.01 level.**	
 E. The intercept is significant at the p<0.01 level.	

```{r}
# Y = B0 + B1X1 + B2X2 + B3X3
# X1 = sqft
# X2 = bedrooms
# X3 = baths

house.main<-lm(price~sqft+bedrooms+baths, data=housing)
summary(house.main)
```


### Question 6 of 19

Explain the effect on price of a one unit increase in sqft in your house.main model.

For every 1 unit increase in square footage, there is a ________ dollar change in price.

```{r}
coef(house.main)
# B sqft = 221.378 -> $221.378 change in price
```


### Question 7 of 19

Build a model, house.inter, that includes the main effect and all interaction terms using sqft, bedrooms, and baths. The response variable is price. 

Which of the following statement(s) are true?

 A. All of the predictors are significant at the p<0.01 level.	
 B. All of the predictors except 1 are significant at the p<0.01 level.	
 C. None of the predictors are significant at the p<0.01 level.	
 D. The intercept is not significant at the p<0.01 level.	
 E. The model is significant at the p<0.01 level.	

```{r}
house.inter<-lm(price~(sqft+bedrooms+baths)^2, data=housing)
summary(house.inter)
```


### Question 8 of 19

Which of the following is the hypothesis for the partial F test to compare the models house.main and house.inter?

 A. Bsqft=Bbedrooms=Bbaths=0
 B. Bsqft=0
 C. Bsqft* bedrooms=Bbedrooms* baths=Bsqft*baths=0
 D. Bsqft=Bbedrooms=BbathsBsqft* bedrooms=Bbedrooms* baths=Bsqft*baths=0


### Question 9 of 19

Conduct the partial F test to compare the models house.main and house.inter. Which of the following statements are true?

 **A. We choose the smaller model.**	
 B. We choose the larger model.	
 C. The models are the same so it doesn't matter which one we choose.	
 **D. The p-value for the partial F-test is greater than 0.05**	
 
```{r}
anova(house.main,house.inter)
```
 


### Question 10 of 19


Use both AIC and adjusted-R2 to compare house.main and house.inter. 

What are the values you find for this comparison?


```{r}
# house.main adjusted-R2: 
summary(house.main)$adj.r.squared

# house.inter adjusted-R2:
summary(house.inter)$adj.r.squared

# house.main AIC: 
AIC(house.main)

# house.inter AIC: 
AIC(house.inter)
```

 

### Question 11 of 19

Based on the AIC and adjusted-R2 values of models house.main and house.inter,  which of the following statements are true?

**A. We choose the smaller model based on AIC and adjusted-R2.**
 B. We choose the larger model based on AIC and adjusted-R2.	
 C. We choose the smaller model based on AIC but the larger model based on adjusted-R2.	
 D. We choose the larger model based on AIC but the smaller model based on adjusted-R2.	
 E. The models are the same based on AIC but the larger model based on adjusted-R2 so we choose the larger model.	


### Question 12 of 19

Plot diagnostics for house.main. How many points have a Cook's distance > 0.5?  ________

```{r}
autoplot(house.main, which=4, ncol = 1, label.size = 3) + theme_bw()
```
```{r}
cooks <- cooks.distance(house.main)
cooks[(cooks > 0.5)] # no points > 0.5
```


### Question 13 of 19

Do you observe any diagnostic problems in the residuals vs. fitted or the q-q plot for house.main model?

 A. Non-constant variance.	
 B. Non-linear fit indicated by non-constant mean.	
 **C. QQ plot shows at least one tail non-Gaussian.**	
 D. No problems observed in diagnostic plots.	
 
```{r}
library(olsrr)
ols_test_breusch_pagan(house.main)
par(mfrow = c(2, 2))
plot(house.main)

autoplot(house.main, which=1, ncol = 1, label.size = 3) + theme_bw()
```
 


### Question 14 of 19

Based on the Box-Cox plot, would you recommend a transformation of the response variable for house.main? If so, what transformation?

 A. No transformation
 **B. Log transformation**
 C. Box cox transformation with L=0.1
 D. Box cox transformation with L=0.5
 
```{r}
boxcox(house.main)
boxcox(house.main, plotit = F)$x[which.max(boxcox(house.main, plotit = F)$y)] 
# log transformation because 0 is in confidence interval
```
 

### Question 15 of 19

If you were to build a model with City as a predictor for Price, which of the following statement(s) would be true? 

 A. Antioch would be the default base case for City.	
 B. Tustin would be the default base case for City.	
 C. This would not be a good idea because there are too many levels of City to use without recoding to fewer levels.	
 D. City does not have a base case	
 E. There are greater than 40 levels of City.	

```{r}
n_distinct(housing$City)

contrasts(as.factor(housing$City))
```


### Question 16 of 19

Using the correlation matrix, find the principal components for the variables price, sqft, bedrooms, and baths. Which 2 variables have the largest absolute loadings in the first principal component?

 A. price	
 *B. sqft*	
 C. bedrooms	
 *D. baths*	

```{r}
house.main.cor <- princomp(housing[,c("price","sqft","bedrooms","baths")], cor = T)
house.main.cor$loadings
```


### Question 17 of 19

How many PCs are needed to explain 90% of the variance in the data?

 A. 1
 B. 2
 C. 3
 D. 4
```{r}
source("PCAplots.R")
cumsum <- cumplot(house.main.cor)
cumsum$plot
```


### Question 18 of 19

Looking at the biplot which variable is most correlated with Price in the first two PCs?

 A. Bedrooms
 B. Baths
 C. Sqft
 D. City
 
```{r}
biplot(house.main.cor, main="Biplot with Correlation Matrix")
```


### Question 19 of 19

Looking at the biplot, which variable is most independent of Price in the first 2 PCs?

 A. Bedrooms
 B. Baths
 C. Sqft
 D. City
