---
title: "Lab 6 Time Series"
author: "Reese Quillian"
date: "2022-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# data and libraries
datadir <- "C:/Users/Student/OneDrive - University of Virginia/Documents/SYS4021/In Class/Data"
sourcedir <- "C:/Users/Student/OneDrive - University of Virginia/Documents/SYS4021"

##load the 'ham_ts.csv' and 'spam_ts.csv' into ham and spam variables repectively
setwd(datadir)
ham<-read.table('ham_ts.csv',header=T,sep=',')
spam<-read.table('spam_ts.csv',header=T,sep=',')
setwd(sourcedir)

##load libraries (install them if necessary)
library(forecast)
library(tidyverse)
library(lubridate)
library(ggplot2)

##use the ts() command to get a time series of ham amount using ham data from all years
ham.ts<-ts(ham$count)

##use the ts() command to get a time series of spam amount using spam data from all years
spam.ts<-ts(spam$count)
```


Question 26 of 48
There are _________ observations in ham_ts.csv and _________ in spam_ts.csv.  
506, 364

```{r}
nrow(ham)
nrow(spam)
```

Question 27 of 48
On average, there are ___________ spam emails per day.
```{r}
mean(spam.ts)
```

Question 28 of 48
On average, there are __________ ham emails per day.  
```{r}
mean(ham.ts)
```

Question 29 of 48
Modify your ham time series object by removing the last six weeks as we discussed in class.  Use this new ham ts object for the remainder of the quiz.

Model the trend of ham using only time as a predictor. Which of the following statements are correct?
 A. There is a significant trend of ham at the 0.05 significance level.	
 B. The coefficient on time in the model is greater than 0.05.	
 C. The number of ham emails decreases with the increase of time.	
 D. There is no significant trend of ham.
 
```{r}
#Remove the last six weeks from ham.ts
ham.ts <-ts(ham$count[1:464])

time.ham<-c(1:(length(ham.ts)))
ham.trend<-lm(ham.ts~time.ham)
summary(ham.trend)
```


Question 30 of 48
Use the acf() function on your ham.ts object. Which of the following statements are true? (select all that apply)
 A. There are no significant autocorrelations in the ham time series after lag=5.	
 B. There are some significant autocorrelations in the ham time series until at least lag=25.	
 C. There are no significant autocorrelations in the acf plot for the ham time series.	
 D. There is a clear seasonality in the ham time series.	
```{r}
acf(ham.ts)
ggAcf(ham.ts)
```
 
 Question 31 of 48
Generate the periodogram for your ham data set. Which of the following statements is true?
 A. There is a cyclical component.
 B. There is a ~7 day period.
 C. There is a monthly seasonal component
 D. There is no obvious seasonality in the ham periodogram.
```{r}
pg.ham <- spec.pgram(ham.ts,spans=9,demean=T,log='no')

# Find the peak, max.omega.ham
max.omega.ham<-pg.ham$freq[which(pg.ham$spec==max(pg.ham$spec))]

# Where is the peak?
max.omega.ham

# What is the period?
1/max.omega.ham
```


Model the trend and the weekly seasonality of ham. Call your new model 'ham.trendseason'. To model the seasonality of ham, create a new categorical predictor variable, Day. Use the days of the weeks as your variables 'F, Sa, S, M, T, W, Th'.
```{r}
Day <- rep(NA, length(ham.ts))
Day[which((time.ham %% 7) == 0)] <- "F"
Day[which((time.ham %% 7) == 1)] <- "Sa"  
Day[which((time.ham %% 7) == 2)] <- "S"
Day[which((time.ham %% 7) == 3)] <- "M"
Day[which((time.ham %% 7) == 4)] <- "T"
Day[which((time.ham %% 7) == 5)] <- "W"
Day[which((time.ham %% 7) == 6)] <- "Th"
Day <- as.factor(Day)

ham.trendseason<-lm(ham.ts~time.ham+Day)
```

Question 32 of 48
Which day of the week is the default base case for your model ?
```{r}
contrasts(Day)
```


Question 33 of 48
Based on your new model, 'ham.trendseason', which of the following statements is true?
 A. Your model ham.trendseason is significant at a 0.05 level.	
 B. None of the days of the week are significantly different than the base case.	
 C. All of the days of the week are significantly different than the base case.	
 D. Only trend is significant.	
```{r}
summary(ham.trendseason)
```

Question 34 of 48
Based on your new model, 'ham.trendseason', which of the following statements is true? (select all that apply)
x A. The frequency of ham e-mail is decreasing on weekdays.	
 B. The frequency of ham e-mail is decreasing on weekends.	
 C. The frequency of ham e-mail on weekdays is not significantly different from the base case.	
 D. None of the above	
```{r}

```




Create a new model of the number of ham emails per day using only the daily seasonality of ham, 'Day', as a predictor. Call your new model 'ham.season'.

```{r}
ham.season<-lm(ham.ts~Day)
```

Question 35 of 48
Is the new model significant at the 0.05 level?
 A. Yes
 B. No
```{r}
summary(ham.season)
```
 
Question 36 of 48
Run a partial F-test between 'ham.trendseason' and 'ham.season'.

Which model do you choose?  Use a 0.05 significance level.
 A. The smaller model.
 B. The larger model.
```{r}
anova(ham.trendseason,ham.season, test="Chi")
```


Question 37 of 48
Use the acf() function on your spam.ts object. Which of the following statements is true?
 A. There are no significant autocorrelations after lag=5.
 B. There are significant autocorrelations after lag=5.
 C. There are no significant autocorrelations in the acf plot for spam.
```{r}
acf(spam.ts)
```

Question 38 of 48
Generate the periodogram for your spam data set. Which of the following statements is true?
 A. There is a weekly seasonal component.
 B. There is a daily seasonal component.
 C. There is a monthly seasonal component.
 D. There is no obvious seasonality in the spam periodogram.
```{r}
pg.spam <- spec.pgram(spam.ts,spans=9,demean=T,log='no')
```

Question 39 of 48.
In your spam periodogram, ωmax is __________ and the period is __________ days.
```{r}
# Find the peak, max.omega.spam
max.omega.spam<-pg.spam$freq[which(pg.spam$spec==max(pg.spam$spec))]

# Where is the peak?
max.omega.spam

# What is the period?
1/max.omega.spam
```

Question 40 of 48
Using only time as a predictor, model the trend of spam for all but the last 7 days of observations (spam.ts[1:(length(spam.ts)-7)]). Call your model 'spam.trend'. Which of the following statements is correct?
 A. There is a significant trend of spam.
 B. The coefficient on time in the model is greater than 0.05.
 C. The number of spam emails decreases with the increase of time.
 D. There is no significant trend of spam.
```{r}
spam.ts <- (spam.ts[1:(length(spam.ts)-7)])

time.spam<-c(1:(length(spam.ts)))
spam.trend<-lm(spam.ts~time.spam)
summary(spam.trend)
```

Question 41 of 48
Get the residuals from the 'spam.trend' model and store them in e.ts.spam. Which of the following statements is true about e.ts.spam?

 A. Its ACF has an exponential decay
 B. Based on its ACF, there are likely autoregressive terms.
 C. There are no autoregressive or moving average terms necessary to model the residuals.
 D. None of the above
```{r}
e.ts.spam <- ts(spam.trend$residuals)
acf(e.ts.spam)
```

Question 43 of 48
Use auto.arima with the argument "approximation=FALSE" on 'e.ts.spam'. Call your new model spam.auto. What is the order of your new model spam.auto?
 A. ARIMA(1,0,0)
 B. ARIMA(1,0,1)
 C. ARIMA(1,1,1)
 D. ARIMA(2,1,0)
```{r}
spam.auto <- auto.arima(e.ts.spam,approximation=FALSE)
summary(spam.auto)
```

Question 44 of 48
The resulting AIC for the model spam.auto is ___________
```{r}
AIC(spam.auto)
```

Question 45 of 48
Fit an ARIMA(1,1,1) to 'e.ts.spam' time series.  Call this new model spam.arima111.

Which of the following statements is true about the non-zero lags of the ACF and PACF plots of the residuals for the spam.arima111 model?
 A. There are no significant spikes in the ACF or PACF.
 B. There are significant spikes in both the ACF and the PACF.
 C. The ACF shows a linear decay.
 D. The PACF has significant spikes from a lag of 5 and beyond.
```{r}
spam.arima111 <- arima(e.ts.spam, order=c(1,1,1))
summary(spam.arima111)


spam.acf <- ggAcf(e.ts.spam)
spam.pacf <- ggPacf(e.ts.spam)

# Plot acf and pacf side by side for easier examination
ggarrange(spam.acf,spam.pacf,nrow=2,ncol=1)

```

Question 46 of 48
Use auto.arima with the argument "approximation=FALSE" to build a new ARIMA model directly to all but the last 7 days of spam.ts (spam.ts[1:(length(spam.ts)-7)]). Call this model spam.auto1.  The ARIMA model has the following order:
p=  ________

d=  ________

q=  _________
```{r}
spam.auto1 <- auto.arima(spam.ts,approximation=FALSE)
summary(spam.auto1)
```

Question 47 of 48
Which of the following statements is true about the non-zero lags of the ACF and PACF plots of the residuals for the spam.auto1 model?
 A. There are no significant spikes in the ACF or PACF.
 B. There are significant spikes in both the ACF and the PACF.
 C. The ACF shows a linear decay.
 D. The PACF has significant spikes from a lag of 5 and beyond.
```{r}
e.ts.spam1 <- ts(spam.auto1$residuals)
spam1.acf <- ggAcf(e.ts.spam1)
spam1.pacf <- ggPacf(e.ts.spam1)

# Plot acf and pacf side by side for easier examination
ggarrange(spam1.acf,spam1.pacf,nrow=2,ncol=1)
```

Question 48 of 48
Which of the following models are adequate based on the Ljung-Box statistic plot? Select all that apply.
 **A. spam.auto up to lag 20.	
 **B. spam.auto1 up to lag 20.	
 **C. spam.arima111 model up to lag 20.	
 D. None of the models are adequate.
```{r}
ggtsdiag(spam.auto,gof.lag=20)

ggtsdiag(spam.auto1,gof.lag=20)

ggtsdiag(spam.arima111,gof.lag=20)
```

