---
title: "Midterm"
author: "Reese Quillian"
date: "2022-10-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, echo = TRUE)
```


```{r}
library(psych)
library(ggplot2)
library(GGally)
library(devtools)
library(ggfortify)
library(here)
library(dplyr)
library(ggpubr)
library(MASS)
library(lindia)
library(olsrr)

setwd("C:/Users/Student/OneDrive - University of Virginia/Documents/SYS4021/")

# read in data
housing_prices <- read.csv("housing-prices.csv")
```

## Part 1

### Question 1

Use a box plot to determine if the price variable has any outliers. If so how many does it have?

```{r}
ggplot(data=housing_prices, aes(x=Price)) + geom_boxplot()
# 7 outliers
```


### Question 2

Use a scatter plot matrix or individual scatter plots to determine the variable in the data set that has the strongest linear relationship with price. What is it?

```{r}
pairs.panels(housing_prices[,c("Rooms", "Baths", "Size", "Age", "Price")])
# size
```


### Question 3

What is the correlation of the strongest linear relationship to Price?

```{r}
# size and price correlation = 0.84
```


### Question 4

Which of the following is a main effects model using Baths as the only predictor variable and y to represent the true value of the observed response, not the modeled prediction?
 A. log(y)= B0+B1X1 + e  where X1 = Baths
 **B. y= B0+B1X1+e  where X1 = Baths**
 C. y= B0+B1X1  where X1 = Baths
 D. log(y)= B0+B1X1 where X1 = Baths

```{r}
# B
```
 

### Question 5

Build a linear model, houses.lm1, to predict Price in terms of Baths. Which of the following is true about the relationship between Baths and Price?

```{r}
houses.lm1<-lm(Price~Baths, data=housing_prices)
summary(houses.lm1)

# B1 is positive -> positive relationship
```


### Question 6

Use your model in Question 5 to explain the effect on price from having one more bathroom. The predicted selling price of a house decreases/increases by ________ dollars per added bathroom. 

```{r}
coef(houses.lm1)
#95298.99
```


### Question 7

Which of the following is a main effects model using all predictor variables (Baths, Rooms, Age, and Size) and y to represent the true value of the observed response, not its modeled prediction?

 **A. y= B0 + B1X1 + B2X2 + B3X3+ B4X4  where X1 = Baths,  X2= Rooms,  X3 = 1 if Old; 0 otherwise,  X4=Size** 
 B. y= B0 + B1X1 + B2X2 + B3X3 + B4X4 + e  where X1 = Baths,  X2 = Rooms,  X3 = 1 if Old; 0 otherwise,  X4 =Size    
 C. y= B0 + B1X1 + B2X2 + B3X3 + B4X4 + B5X1X2 + B6X1X3 + B7X1X4 + B8X2X3 + B9X2X4 + B10X3X4  where X1 = Baths,  X2 = Rooms,  X3 = 1 if Old; 0 otherwise,  X4 = Size  
 D. y= B0+B1X1 + B2X2 + B3X3 + B4X4 + B5X1X2 + B6X1X3 + B7X1X4 + B8X2X3 + B9X2X4 + B10X3X4 + e  where X1 = Baths,  X2= Rooms,  X3 = 1 if Old; 0 otherwise,  X4 = Size    

### Question 8

Which of the following is a model that includes the main effect and all interaction terms using ONLY Size and Age as predictors of the response, whose true, observed value is represented by y.

 A. y= B0+ B1X1 + B2X2  where X1 = Size,  X2 = 1 if Old; 0 otherwise 
 B. y= B0 + B1X1 + B2X2 + e  where X1 = Size,  X2 = 1 if Old; 0 otherwise  
 **C. y= B0 + B1X1 + B2X2 + B3X1X2  where X1 = Size,  X2 = 1 if Old; 0 otherwise**
 D. y= B0 + B1X1 + B2X2 + B3X1X2 + e  where X1 = Size  X2 = 1 if Old; 0 otherwise  



### Question 9
 
How many parameters does your interaction model in Question 8 have? **4**

Coefficients?  **3**


### Question 10

What is the null hypothesis for the Partial F-test to compare the models you found in Questions 4 and 7?

 A. B1=0
 B. B1=B2=0
 **C. B2=B3=B4=0**
 D. B4=0
 

### Question 11

Conduct the partial F test to compare the models you found for Questions 4 and 7.  What is the exact p-value (not the significance level)?

```{r}
houses.lm2<-lm(Price~Baths+Rooms+Age+Size, data=housing_prices)
summary(houses.lm2)

anova(houses.lm1, houses.lm2)
```



### Question 12

What do the results of the partial F test comparing the models you found for Questions 4 and 7 suggest with regards to which model you should choose?

 **A. Reject the null hypothesis and therefore use the larger model with additional terms.**
 B. Fail to reject this hypothesis and therefore use the smaller model without additional terms.
 C. Reject the null hypothesis and therefore use the smaller model without additional terms.
 D. Fail to reject this hypothesis and therefore use the larger model with additional terms.


### Question 13
 
Now create a stepwise model from a main effects + interaction model for ALL predictor variables.  What is the AIC of the model? Round to the nearest whole number.
 
```{r}
houses.inter <- lm(Price~(Baths+Rooms+Age+Size)^2,data=housing_prices)
#summary(houses.inter)

#step
houses.step<-step(houses.inter, trace=T)
summary(houses.step)
#AIC(houses.step)
```
 


### Question 14

For your stepwise model from Question 13 (a main effects + interaction model for ALL predictor variables).  How many parameters are significant at the 0.05 level? 

```{r}
# 7 significant paramaeters (includes intercept)
```
 

### Question 15

Model diagnostics

```{r}
par(mfrow = c(2, 2))
plot(houses.step)

# non-constant variance
ols_test_breusch_pagan(houses.step) # p-value <.05 so there is non-constant variance

autoplot(houses.step, which=4, ncol = 1, label.size = 3) + theme_bw()
```


### Question 16

For your stepwise model from Question 13, using Box-Cox plot would you recommend a transformation of the response variable and if so, what type?

```{r}
boxcox(houses.step)
boxcox(houses.step, plotit = F)$x[which.max(boxcox(houses.step, plotit = F)$y)]

# power transformation with L=0.5
```


### Question 17

Answer the following questions about the point with the highest Cook's distance.

```{r}
# observation number 9
# other info:
housing_prices[9,]
```


### Question 18

Which variables could not be included in principal components analysis of this dataset?
- anything quantitative (i.e. NOT AGE)


### Question 19

Using the correlation matrix, find the principal components for the dataset using all variables that can be included in PCA.  Which 2 variables have the largest absolute loadings in the first principal component?

```{r}
houses.corr <- princomp(housing_prices[,c("Price","Rooms","Baths","Size")], cor = T)
houses.corr$loadings

# size and price
```

### Question 20

In the first PC, Price moves in the same direction as:

rooms, baths, and size because they are all positive


### Question 21

Which 2 variables have the largest absolute loadings in the second principal component? 
rooms and price

### Question 22

In the second PC, Price moves in the same direction as:
size only

### Question 23

Which of the following is true about the scree plot

```{r}
# scree plot
source("PCAplots.R")
ggscreeplot(houses.corr)
```


### Question 24

How many principal components would it take to account for 90 percent of the variance?

```{r}
cumplot(houses.corr)
# 3 pcs
```

### Question 25

Make a biplot of the data in the first two PCs. Which variables appear most correlated in the first two PCs?

```{r}
biplot(houses.corr)
# price and size
```

### Question 26

Based on the biplot, which two variables appear most independent in the first two PCs?

closest to perpendicular -> price and rooms

### Question 27

Find the observation with the largest absolute value of PC2. What is the size of this house in square feet?

```{r}
ggbiplot(houses.corr, varname.size = 5, labels=row(housing_prices)[,1], plot.obs=TRUE, 
         xlim=c(-0.2,0.4), ylim=c(-0.3,0.3))

# observation 88:

housing_prices[88,]
#housing_prices[2,]
```

## Part 2

### Question 28

Make a boxplot of child_mort. How many outliers are there? (Hint: use the statistics of the boxplot if you can't count them in the figure)   

```{r}
country <- read.csv("Country-data.csv")
```

```{r}
ggplot(data=country, aes(x=child_mort)) + geom_boxplot()
boxplot.stats(country$child_mort)
```


### Question 29

What countries do the outliers correspond to?

```{r}
# observations are where child mort = 149, 150, 208, 160
country %>% filter(child_mort %in% c(149,150,208,160))
```

### Question 30

Which country has the lowest child mortality rate?

```{r}
country %>% arrange(child_mort) %>% head()
```

### Question 31

Make a scatter plot of all variables in the dataset except "Country." Which variable's correlation coefficient with child_mort has the greatest absolute value?

```{r}
pairs.panels(country[,c("child_mort", "exports", "health", "imports", "income","inflation","life_expec","total_fer","gdpp")])
```

### Question 32

Using the correlation matrix, perform PCA on this dataset, excluding the "Country" variable. How many PCs are needed to explain 90% of the variance in the dataset? 

```{r}
countries.corr <- princomp(country[,c("child_mort", "exports", "health", "imports", "income","inflation","life_expec","total_fer","gdpp")], cor = T)
```

```{r}
cumplot(countries.corr)
```

### Question 33
Make a biplot of the data in the first two PCs. Which variable is most positively correlated with child mortality in the first two PCs?

```{r}
biplot(countries.corr)
# without observations
ggbiplot(countries.corr, varname.size = 3, labels=row(country)[,1], plot.obs=TRUE, 
         xlim=c(-3,3), ylim=c(-3,3))
```

### Question 34

Based on the biplot, which variable is most unrelated to child_mort?
imports, almost exactly 90 degrees

### Question 35

```{r}
# obs 114
country[114,]
```

### Question 36
```{r}
# highest in PC 2 is obs 134
country[134,]
```

### Question 37

How many variables vary in the same direction as child_mort in the first PC?
```{r}
countries.corr$loadings
```

### Question 38

Which two variables have the largest absolute loading in the second PC?

exports and imports